### 2021-05-11 article/article.tex
update

The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux at detector level for flavor $\beta$ as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ on both probability
and flux level, we plot the oscillograms resulting from Eq.~\ref{eq:propFlux} in Fig.~\ref{fig:flux_ratio}. In the left panel, we have marked the region in which 99\% of the 
DeepCore cascade events originating from $\ne$ and $\nt$ fluxes are contained. In the right panel, we show the two regions in which 99\% of the IceCube 
and DeepCore track events originating from $\nm$ fluxes are contained. Starting with the $\nm$ flux ratio, we see that the only clear signal discernible to the IceCube detector
is a flux deficiency of a factor of $\sim 10^2$ from core-crossing neutrinos within a zenith range of $\ztrue > -0.87$. DeepCore on the other hand, 
observes multiple fringes of flux surpluses with a factor $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensetive IceCube deficiency.

For the fluxes which drives cascades, we have to resort to the DeepCore detector alone. %TODO: fill more
Here we see a somewhat weaker signal, this time a zenith-independent deficiency. 
%TODO: fill with conclusion

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.35\linewidth]{figures/cascade_flux_ratio.pdf}
      \hspace{1.5cm}
      \includegraphics[width=0.35\linewidth]{figures/track_flux_ratio.pdf}
   \end{center}
   \caption{Ratio of propagated NSI to SI atmospheric fluxes at detector level. We set $\emt = -0.05$, and all other NSI parameters to zero. 
   Dotted (dashed) lines show the regions in which 99\% of the 
   DeepCore (IceCube) MC events are contained. \emph{Left panel:} Propagated fluxes of $\ne$ and $\nt$ neutrinos and anti-neutrinos. 
   \emph{Right panel:} Propagated fluxes of $\nm$ neutrinos and anti-neutrinos}\label{fig:flux_ratio}%TODO: normalize color scale and fix limits
\end{figure}% smaller emt and no log. clarify that IC is missing in left panel. clarify
%that energy region within dotted line, use as few words as possible. Normalize color scale. check green-yellow cmap

\subsection{IceCube}\label{ch:ICmethod}
The event rate for each bin reads
\begin{align}\label{eq:ICevents}
   N_{ij} &= T\sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty \phi_\beta^\text{det}  A^\text{eff} R(E^r,E^t) 
   \dd E^t\,,
\end{align}
where $T$ is the live time of the detector, and $A^\text{eff}$ its effective area averaged over the flavors $\beta$ from~\cite{ICaeff}. $R(x^r,x^t)$ is a resolution function, 
which is responsible for the smearing between the reconstructed and true parameters $x^r$ and $x^t$, respectively. We assume a log-normal distribution, giving it the form 
\begin{align}
    R(x^r, x^t) = \frac{1}{\sqrt{2\pi} \sigma_{x^r}x^r} \exp\left[-\frac{(\log x^r-\mu(x^t))^2}{2\sigma_{x^r}^2}\right]\,.
\end{align}
As seen in Fig.~\ref{fig:IC_MC_counts}, the energy reconstruction is biased. To model this relationship between $\Etrue$ and $\Ereco$, we train a Gaussian process regressor on the dataset~\cite{IC2016}, from which
we can extract a predicted mean and standard deviation for a given $E^{reco}$. We then take the $\Etrue$ points of the 99th percentile of each distribution to obtain
the limits of $\Etrue$ at which to integrate over. We have no angular resolution function since the angle resolution in IceCube for track-like events is less than $\SI{2}{\degree}$, making $\ztrue$ coincide with $\zreco$ for our study. 
The data is from the 86 string sterile analysis~\cite{IC2020}.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{SI}$ $(N_{NSI})$ are the numbers of expected events
assuming standard (non-standard) interactions in Fig.~\ref{fig:DC_event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.

\begin{figure}[t]
   \begin{center}
         \includegraphics[width=0.9\textwidth]{figures/event_pulls.pdf}
   \end{center}
   \caption{Expected pulls of the form $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ for PINGU and DeepCore after 3 years. We compare the NSI event count with $\emt=-0.01$
    to the standard interaction count}\label{fig:event_pulls} 
\end{figure}

\subsection{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations are the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU MC is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. Just as with DeepCore, 
we plot the events pulls for cascades and tracks in Fig.~\ref{fig:PINGU_event_pulls}. %TODO:talk about this
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

@@

\begin{figure}[!tb]
   \begin{center}
      \begin{subfigure}{0.4\textwidth}
         \includegraphics[width=1\linewidth]{figures/PINGU_3D_NO.pdf}
         \caption{PINGU}\label{fig:PINGU_3D_NO}
      \end{subfigure}
      \hspace{1cm}
      \begin{subfigure}{0.4\textwidth}
         \includegraphics[width=1\linewidth]{figures/DC_3D_NO.pdf}
         \caption{DeepCore}\label{fig:DC_3D_NO}
      \end{subfigure}
    \end{center}
    \caption{Minimization of the $\Delta \chi^2$ defined in Eq.~\ref{eq:chisq} after three years of data.
    $\dm$ and $\theta_{23}$ and have been marginalized out, and all other NSI 
    parameters other than the one shown in each panel are fixed to zero.}\label{fig:3D_NO}
 \end{figure}
%TODO:explain emt assymetry. dont put pingu and dc side by side since pingu is generated.. include 3sigma if possible?
%TODO: ploit contour for ett-emt
 \begin{table}
   \begin{center}
   \begin{tabular}{lcc}
           \hline \hline \\
           Parameter & 90\% CL & $3\sigma$  \\
           \hline \multicolumn{3}{c} {\hspace{1.5cm} DeepCore }  \\[0.1em]
           $\ett$ & [-0.029, 0.044]\hspace{1cm} & [-, -] \\
           $\emt$ & [-0.015, 0.005]\hspace{1cm} & [-, 0.015] \\
           $\eem$ & [-0.072, 0.075]\hspace{1cm} & [-0.132, 0.127]\\
           $\eet$ & [-0.074, 0.141]\hspace{1cm} & [-0.174, 0.2  ]\\\\
           \multicolumn{3}{c} {\hspace{1.5cm} PINGU } \\
           $\ett$ & [-0.033, -]\hspace{1cm} & [-0.048, -] \\
           $\emt$ & [-0.007, 0.007]\hspace{1cm} & [-0.011, 0.013] \\
           $\eem$ & [-0.062, 0.057]\hspace{1cm} & [-0.122, 0.103] \\
           $\eet$ & [-0.069, 0.077]\hspace{1cm} & [-0.121, 0.133] \\\\
           \multicolumn{3}{c} {\hspace{1.5cm} DeepCore + PINGU } \\
           $\ett$ & [-0.028, 0.043]\hspace{1cm} & [-0.043, -]\\
           $\emt$ & [-0.007, 0.005]\hspace{1cm} & [-0.011, 0.01] \\
           $\eem$ & [-0.048, 0.048] \hspace{1cm} & [-0.09, 0.085]\\
           $\eet$ & [-0.054, 0.072] \hspace{1cm} & [-0.101, 0.118] \\
           \hline
           \hline
   \end{tabular}
   \end{center}
   \caption{DeepCore and PINGU results from the $\Delta \chi^2$ in Fig.~\ref{fig:3D_NO}.}\label{table:DC_PINGU_results}
\end{table}

@@

%TODO: segue to joint
For the joint analysis, we follow the parameter goodness-of-fit prescription~\cite{maltoni2003} and construct the joint $\chi^2$ as 
\begin{align}\label{eq:joint_chisq}
    \chi^2_\text{joint} = \chi^2_\text{IC} + \chi^2_\text{DC} + \chi^2_\text{P} - \chi^2_{IC,min} - \chi^2_\text{DC,min} - \chi^2_\text{P,min}\,
\end{align}
with test statistic $\chi^2_\text{joint,min}$. The results are shown in Fig.~\ref{fig:joint_3D} and summarized in Table~\ref{table:DC_PINGU_results}.

\begin{figure}[t]
    \begin{center}
       \includegraphics[width=0.9\linewidth]{figures/joint_3D_NO.pdf}
    \end{center}
    \caption{Expected joint $\Delta \chi^2$ from Eq.~\ref{eq:joint_chisq} for normal ordering.}
    \label{fig:joint_3D}
 \end{figure}

%TODO:  Discussion on impact of IC on chisq.?

>>>>

The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux at detector level for flavor $\beta$ as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ on both probability
and flux level, we plot the oscillograms resulting from Eq.~\ref{eq:propFlux} in Fig.~\ref{fig:flux_ratio}. In the left panel, we have marked the region in which 99\% of the 
DeepCore cascade events originating from $\ne$ and $\nt$ fluxes are contained. In the right panel, we show the two regions in which 99\% of the IceCube 
and DeepCore track events originating from $\nm$ fluxes are contained. Starting with the $\nm$ flux ratio, we see that the only clear signal discernible to the IceCube detector
is a flux deficiency of a factor of $\sim 10^2$ from core-crossing neutrinos within a zenith range of $\ztrue > -0.87$. DeepCore on the other hand, 
observes multiple fringes of flux surpluses with a factor $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensetive IceCube deficiency.
%TODO: check values above after final plot of flux ratios
For the fluxes which drives cascades, we have to resort to the DeepCore detector alone. %TODO: fill more
Here we see a somewhat weaker signal, this time a zenith-independent deficiency. 
%TODO: fill with conclusion

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.35\linewidth]{figures/cascade_flux_ratio.pdf}
      \hspace{1.5cm}
      \includegraphics[width=0.35\linewidth]{figures/track_flux_ratio.pdf}
   \end{center}
   \caption{Ratio of detector NSI to SI atmospheric fluxes. We set $\emt = -0.05$, and all other $\epsilon_{\alpha\beta}=0$. 
   Within the dotted (dashed) lines, 99\% of the  DeepCore (IceCube) MC events are contained. Since we limit our IceCube study to muon tracks,
   a dashed region is omitted in the left panel.}
\end{figure}%Done, run code in nsi.ipynb. need high emt and log to show both cascade and track features at the same time.
%otherwise, the cascade ratio overpowers the track one.

\subsection{IceCube}\label{ch:ICmethod}
The event rate for each bin reads
\begin{align}\label{eq:ICevents}
   N_{ij} &= T\sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty \phi_\beta^\text{det}  A^\text{eff} R(E^r,E^t) 
   \dd E^t\,,
\end{align}
where $T$ is the live time of the detector, and $A^\text{eff}$ its effective area averaged over the flavors $\beta$ from~\cite{ICaeff}. $R(x^r,x^t)$ is a resolution function, 
which is responsible for the smearing between the reconstructed and true parameters $x^r$ and $x^t$, respectively. We assume a log-normal distribution, giving it the form 
\begin{align}
    R(x^r, x^t) = \frac{1}{\sqrt{2\pi} \sigma_{x^r}x^r} \exp\left[-\frac{(\log x^r-\mu(x^t))^2}{2\sigma_{x^r}^2}\right]\,.
\end{align}
However, the energy reconstruction is biased~\cite{weaverEvidenceAstrophysicalMuon}, which means that we don't make the assumption that $\mu(E^t) =E^r$. To model this relationship between $\Etrue$ and $\Ereco$, we train a Gaussian process regressor on the dataset~\cite{IC2016}, from which
we can extract a predicted mean and standard deviation for a given $E^{reco}$. We then take the $\Etrue$ points of the 99th percentile of each distribution to obtain
the limits of $\Etrue$ at which to integrate over. We have no angular resolution function since the angle resolution in IceCube for track-like events is less than $\SI{2}{\degree}$, making $\ztrue$ coincide with $\zreco$ for our study. 
The data is from the 86 string sterile analysis~\cite{IC2020}.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{SI}$ $(N_{NSI})$ are the numbers of expected events
assuming standard (non-standard) interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.

\subsection{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations are the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU MC is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. Just as with DeepCore, 
we plot the events pulls for cascades and tracks in Fig.~\ref{fig:event_pulls}. %TODO:talk about this
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

@@

\begin{figure}[!tb]
   \begin{center}
      \begin{subfigure}{0.45\textwidth}
         \includegraphics[width=1\linewidth]{figures/joint_3D_NO.pdf}
         \caption{Normal ordering}\label{fig:joint_3D}
      \end{subfigure}
      \hspace{0cm}
      \begin{subfigure}{0.4\textwidth}
         \includegraphics[width=1\linewidth]{figures/DC_3D_NO.pdf}
         \caption{DeepCore}\label{fig:DC_3D_NO}
      \end{subfigure}
    \end{center}
    \caption{Expected $\Delta \chi^2$ after three years of data.
    $\dm$ and $\theta_{23}$ and have been marginalized out, and all other NSI 
    parameters other than the one shown in each panel are fixed to zero.}\label{fig:3D_NO}
 \end{figure}
%TODO:explain emt assymetry. dont put pingu and dc side by side since pingu is generated.. include 3sigma if possible?
%TODO: ploit contour for ett-emt
 \begin{table}
   \begin{center}
   \begin{tabular}{lcc}
           \hline \hline \\
           Parameter & 90\% CL & $3\sigma$  \\
           \hline \multicolumn{3}{c} {\hspace{1.5cm} DeepCore }  \\[0.1em]
           $\ett$ & [-0.029, 0.044]\hspace{1cm} & [-, -] \\
           $\emt$ & [-0.015, 0.005]\hspace{1cm} & [-, 0.015] \\
           $\eem$ & [-0.072, 0.075]\hspace{1cm} & [-0.132, 0.127]\\
           $\eet$ & [-0.074, 0.141]\hspace{1cm} & [-0.174, 0.2  ]\\\\
           \multicolumn{3}{c} {\hspace{1.5cm} PINGU } \\
           $\ett$ & [-0.033, -]\hspace{1cm} & [-0.048, -] \\
           $\emt$ & [-0.007, 0.007]\hspace{1cm} & [-0.011, 0.013] \\
           $\eem$ & [-0.062, 0.057]\hspace{1cm} & [-0.122, 0.103] \\
           $\eet$ & [-0.069, 0.077]\hspace{1cm} & [-0.121, 0.133] \\\\
           \multicolumn{3}{c} {\hspace{1.5cm} DeepCore + PINGU } \\
           $\ett$ & [-0.028, 0.043]\hspace{1cm} & [-0.043, -]\\
           $\emt$ & [-0.007, 0.005]\hspace{1cm} & [-0.011, 0.01] \\
           $\eem$ & [-0.048, 0.048] \hspace{1cm} & [-0.09, 0.085]\\
           $\eet$ & [-0.054, 0.072] \hspace{1cm} & [-0.101, 0.118] \\
           \hline
           \hline
   \end{tabular}
   \end{center}
   \caption{DeepCore and PINGU results from the $\Delta \chi^2$ in Fig.~\ref{fig:3D_NO}.}\label{table:DC_PINGU_results}
\end{table}

@@

%TODO: segue to joint
For the joint analysis, we follow the parameter goodness-of-fit prescription~\cite{maltoni2003} and construct the joint $\chi^2$ as 
\begin{align}\label{eq:joint_chisq}
    \chi^2_\text{joint} = \chi^2_\text{IC} + \chi^2_\text{DC} + \chi^2_\text{P} - \chi^2_{IC,min} - \chi^2_\text{DC,min} - \chi^2_\text{P,min}\,
\end{align}
with test statistic $\chi^2_\text{joint,min}$. The results are shown in Fig.~\ref{fig:joint_3D} and summarized in Table~\ref{table:DC_PINGU_results}.
\newpage

 \begin{figure}[t]
   \begin{center}
      \begin{subfigure}{0.38\textwidth}
         \includegraphics[width=1\linewidth]{figures/PINGU_event_pulls.pdf}
         \caption{PINGU}\label{fig:PINGU_event_pulls}
      \end{subfigure}
      \begin{subfigure}{0.4\textwidth}
         \includegraphics[width=1\linewidth]{figures/DC_event_pulls.pdf}
         \caption{DeepCore}\label{fig:DC_event_pulls}
      \end{subfigure}
    \end{center}
   \caption{Expected pulls of the form $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ for PINGU and DeepCore after 3 years. We compare the NSI event count with $\emt=-0.01$
    to the standard interaction count}\label{fig:event_pulls} 
\end{figure}%TODO: redo these with same emt as flux_ratio?

% \begin{figure}[b]
%    \begin{center}
%       \includegraphics[width=0.2\linewidth]{figures/joint_3D_NO.pdf}
%    \end{center}
%    \caption{Expected joint $\Delta \chi^2$ from Eq.~\ref{eq:joint_chisq} for normal ordering.}
%    \label{fig:joint_3D}
% \end{figure}
%TODO:  Discussion on impact of IC on chisq.?
